---
title: 'Stat 480 - Homework #9'
author: "Zahra Khoshmanesh"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


#### Weekly box office data

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.

2. For this homework we use the data set `box` from the `classdata` package, which consists of weekly box office gross for movies of the last five year.

```{r, message=FALSE, warning=FALSE}
# devtools::install_github("haleyjeppson/classdata")
library(classdata)
library(tidyverse)

head(box)
```

3. In class we discussed two instances where a movie was released under the same name as a different movie previously. Identify at least one more instance of a movie where that happened. Report the name of the movie, search online for additional information. Describe the strategy you used to identify this movie, report the code involved.  

```{r}
## your answer here
library(lubridate)

distinct(box) %>%
  mutate(Year=year(Date)) %>%
  group_by(Movie) %>%
  summarise(count = length(unique(Year))) %>% filter(count > 1)

```

** Describe**
I grouped the movies by their names and count the unique number of years that movie has record, then filter the movies has more than 1 unique yearsince it indicates the movie remake under same name later.
Let us know more about Daddy’s Home .

```{r}
box %>% filter(Movie=="Daddy’s Home")

box %>% filter(Movie=="Daddy’s Home 2")
```



4. Re-derive variables: `Change` (percent change in gross income from last week), `Rank.Last.Week`, `Per.Thtr.` (as gross per theater), and `Total.Gross` (as the cumulative sum of weekly gross).

```{r}
## your answer here

# Re-derive variables: Change (percent change in gross income from last week), Rank.Last.Week, 
# Per.Thtr. (as gross per theater), and Total.Gross (as the cumulative sum of weekly gross).

redrive <- box %>% 
  group_by(Movie,Thtrs.) %>% 
  mutate(Per.Thtr.=sum(Gross))

sample <- box %>% 
  filter(Movie == "13 Minutes") %>% 
  arrange(Week)

lag()
lag(1:10, 1)



sample %>% 
  mutate(rank_last_week = lag(Rank, order_by = Week)) %>% 
  select(Week, Rank, Rank.Last.Week, rank_last_week)

sample %>% 
  mutate(week_diff = Week - lag(Week),
         rank_last_week = lag(Rank, order_by = Week)) %>% 
  select(Week, Rank, Rank.Last.Week, week_diff, rank_last_week)

# ifelse example
x <- c(6:-4)
sqrt(x)  #- gives warning
ifelse(x >= 0, x, NA)
sqrt(ifelse(x >= 0, x, NA))  # no warning

sample %>% 
  mutate(week_diff = Week - lag(Week),
         rank_last_week = ifelse(week_diff == 1, lag(Rank, order_by = Week), NA)) %>% 
  select(Week, Rank, Rank.Last.Week, -week_diff, rank_last_week)
```


5. For the variables `Per.Thtr.` and `Change` compare the original variables and the newly derived ones. Are there differences? Where? Try to describe patterns you find.

```{r}
## your answer here
```


6. Is the original variable `Total.Gross` strictly increasing? 

```{r}
## your answer here
```


7. Identify the three top grossing movies for each year. 
Plan of attack: 
    - Extract the year from the `Date` variable.
    - Summarize the total gross for each movie and each year.
    - Find the rank of movies by total gross in each year.
    - Filter the top three movies.


```{r}
## your answer here
```

