---
title: 'Stat 480 - Homework #2'
author: "Zahra Khoshmansh"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Economic Guide to Picking a College Major

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.

2. Using the command below, read in the data set into your R session. 

```{r}
recent_grads <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/materials-2020/master/02-r_intro/data/recent_grads.csv")

```

3. Create a new variable, `share_women`, that is women as share of total; i.e. the number of women divided by the total number of men and women.

4. Create a subset of the data containing only the rows where the `major_category` is `STEM`.

5. For your subset, compute the average  of `share_women` and its standard deviation. Also compute the average median earnings (`median`) and its standard deviation. Comment on the results. (You might have to deal with missing values appropriately).

5. Compute the correlation between women as a share of total (`share_women`) and the median earnings (`median`) and interpret your results.

6. Use the original dataset and `ggplot2` to draw a scatterplot of women as share of total and the median earnings. Color points by the major category (`major_category`). Comment on the result. 

Due date: the homework is due before class on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html file with it. 


## Solution

```{r}
#2-read dataset
recent_grads <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/materials-2020/master/02-r_intro/data/recent_grads.csv")

```

```{r}
#3-Create a new variable, `share_women`, that is women as share of total; i.e. the number of women divided by the total number of men and women.
recent_grads$share_women <- recent_grads$Women/(recent_grads$Women + recent_grads$Men)
head(recent_grads)
```


```{r}
#4-Create a subset of the data containing only the rows where the `major_category` is `STEM`
STEM <- recent_grads %>% filter(Major_category %in% "STEM")
head(STEM)

```


```{r}
#5-For your subset, compute the average  of `share_women` and its standard deviation. Also compute the average median earnings (`median`) and its standard deviation. Comment on the results. (You might have to deal with missing values appropriately).
mean_share_woman <- mean(recent_grads$share_women,na.rm = TRUE)
mean_share_woman
sd_share_woman <- sd(recent_grads$share_women,na.rm = TRUE)
sd_share_woman

mean_median_earnings <- mean(recent_grads$Median,na.rm = TRUE)
mean_median_earnings
sd_median_earnings <- sd(recent_grads$Median,na.rm = TRUE)
sd_median_earnings
```



```{r}
#6- Compute the correlation between women as a share of total (`share_women`) and the median earnings (`median`) and interpret your results.

correlation <- cor.test(recent_grads$share_women,recent_grads$Median, method = "pearson")
correlation

```

correlation is -0.6186898 which indicates variables which can be considered moderately correlated. 

```{r}
#7- Use the original dataset and `ggplot2` to draw a scatterplot of women as share of total and the median earnings. Color points by the major category (`major_category`). Comment on the result.
ggplot(data= recent_grads,aes(x=share_women,y=Median))+
  geom_point(aes(color=Major_category))

ggplot(data = recent_grads, aes(x = log(share_women), y = log(Median))) +
  geom_point(aes(color=Major_category))


  

```


Interpretation of result:

1. Form: Linear
2. Strength: not strong. Data points are scattered.
3. Direction: Negatively Associated.
4. an outlier. look at below code

```{r}
#check for outlier in indivisual variables
boxplot(recent_grads$share_women)
boxplot(recent_grads$Median)#median has some outliers

# use cook distance to check the outlier
mod <- lm(share_women ~ recent_grads$Median, data=recent_grads)
cooksd <- cooks.distance(mod)

plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels

```

