---
title: 'Stat 480 - Homework #5'
author: "Zahra Khoshmanesh"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```


## Flying etiquette

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.


2. FiveThirtyEight is a website founded by Statistician and writer Nate Silver to publish results from  opinion poll analysis, politics, economics, and sports blogging. One of the featured articles considers [flying etiquette](https://fivethirtyeight.com/features/airplane-etiquette-recline-seat/). This article is based on data collected by FiveThirtyEight and publicly available on github. Use the code below to read in the data from the survey:

```{r}
fly <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/flying-etiquette-survey/flying-etiquette.csv")
```

The next couple of lines of code provide a bit of cleanup of the demographic information by reordering the levels of the  corresponding factor variables. Run this code in your session.

```{r}
fly$Age <- factor(fly$Age, levels=c("18-29", "30-44", "45-60", "> 60", ""))
fly$Household.Income <- factor(fly$Household.Income, levels = c("$0 - $24,999","$25,000 - $49,999", "$50,000 - $99,999", "$100,000 - $149,999", "150000", ""))
fly$Education <- factor(fly$Education, levels = c("Less than high school degree", "High school degree", "Some college or Associate degree", "Bachelor degree",  "Graduate degree", ""))
```


3. Some people do not travel often by plane. Provide a breakdown of travel frequency (use variable `How.often.do.you.travel.by.plane.`). Reorder the levels in the variable by travel frequency from least frequent travel to most frequent. Draw a barchart of travel frequency and comment on it. 

```{r}
#factor levels befor reordering
levels(fly$How.often.do.you.travel.by.plane.)
#check if we have empty value for this variable
any(fly$How.often.do.you.travel.by.plane.==" ")
#reorder
fly$How.often.do.you.travel.by.plane. <- factor(fly$How.often.do.you.travel.by.plane.,
                                                levels=c("Never","Once a year or less","Once a month or less" ,"A few times per month",
                                                         "A few times per week","Every day"))
#factor levels after reordering
levels(fly$How.often.do.you.travel.by.plane.)


```

If question is to order levels based on the frequency of levels in the data :

most frequent level ordered as follows :

"Once a year or less"
"Once a month or less"
"Never"
"A few times per month"
"A few times per week"
"Every day"

```{r}
# if question ask to reorder levels based on the frequency of levels in data

fly$How.often.do.you.travel.by.plane. <- factor(fly$How.often.do.you.travel.by.plane.,
                                                levels=c("Once a year or less",
                                                         "Once a month or less",
                                                         "Never",
                                                         "A few times per month",
                                                         "A few times per week",
                                                         "Every day"))

```

Draw a barchart of travel frequency and comment on it

```{r}
library(forcats)

ggplot(data=fly, aes(x=fct_infreq(fly$How.often.do.you.travel.by.plane.),fill=fly$How.often.do.you.travel.by.plane.)) +
  geom_bar()+ 
  coord_flip()

```

**Chart Interpretation**

based on the result of abov egraph:
most of people travel once a year or less following with once a month or less since these two categories are the highest frequent categories.


4. In the demographic variables (`Education`, `Age`, and `Houshold.Income`), replace all occurrences of the empty string "" by a missing value `NA`. How many responses do not have any missing values? (Hint: the function `is.na` might come in handy)

```{r}
#let us check the na value in education,age and household.income
str(fly$Education)
levels(fly$Education)
levels(fly$Education)[6] <- NA
levels(fly$Education)


str(fly$Age)
levels(fly$Age)
levels(fly$Age)[5] <- NA
levels(fly$Age)

str(fly$Household.Income)
levels(fly$Household.Income)
levels(fly$Household.Income)[6] <- NA
levels(fly$Household.Income)



```

How many responses do not have any missing values? (Hint: the function `is.na` might come in handy)


let us first calculate how many responds has the missing value in one of these three variables: age, education, household.income

```{r}
sum((is.na(fly$Age) | is.na(fly$Education) | is.na(fly$Household.Income)))
```
 216 rows has missing value in at least one of these three variables.
 
 now we use logical operator ! and count the number of rows that do not have missing value.
 
```{r}
sum(!(is.na(fly$Age) | is.na(fly$Education) | is.na(fly$Household.Income)))

```
and 824 rows do not have missing value in these three variables.


5. Run the command below and interpret the output. What potential purpose can you see for the chart? 
```{r, fig.width=8, fig.height=3, message=FALSE}
library(ggplot2)
fly$Education = with(fly, factor(Education, levels = rev(levels(Education))))

ggplot(data = fly, aes(x = 1)) + 
  geom_bar(aes(fill=Education), position="fill") + 
  coord_flip() +
  theme(legend.position="bottom") +
  scale_fill_brewer() + 
  xlab("Ratio") 
```

**INterpretation**
first it reverse the level of education factor to start from less than highschool to graduate colledge. Then use the ratio of frequency of each levels and draw a diagram. the sum of all ratios are 1. the width of each level in the chart shows the ratio. the highest ratio belongs to bachelor degree.
this chart shows the ratio of  education levels. as we can see by seeing the width of each color, the bachelor degree has the highest ratio following it the Some college or Associate degree and graduate degree. the NA has the lowest fregency in the data meaning very small ratio.

```{r}
#calculate the ratio to confirm the result of interpretation
fly %>% select(Education) %>% group_by(Education) %>% summarise(ratio=n()/nrow(fly))
```



6. Rename the variable `In.general..is.itrude.to.bring.a.baby.on.a.plane.` to `baby.on.plane.`. How many levels does the variable `baby.on.plane` have, and what are these levels? Rename the level labeled "" to "Not answered". Reorder the levels of `baby.on.plane` from least rude to most rude. Put the level "Not answered" last. Draw a barchart of variable `baby.on.plane`.  Interpret the result.

```{r}
names(fly) #check the index of In.general..is.itrude.to.bring.a.baby.on.a.plane.
names(fly)[19] <- "baby.on.plane."
length(levels(fly$baby.on.plane.))
levels(fly$baby.on.plane.)

```
 it has 4 levels : 
 ""   
 "No, not at all rude" 
 "Yes, somewhat rude" 
 "Yes, very rude" 
 
```{r}
levels(fly$baby.on.plane.)[1] <- "Not answered"
fly$baby.on.plane. <- factor(fly$baby.on.plane.,levels=c("No, not at all rude","Yes, somewhat rude" ,"Yes, very rude","Not answered"))
levels(fly$baby.on.plane.)

```
 
Draw a barchart of variable `baby.on.plane`.  Interpret the result.

```{r}
ggplot(data=fly, aes(x=fct_infreq(fly$baby.on.plane.),fill=fly$baby.on.plane.)) +
  geom_bar()+ 
  coord_flip()

```

**Interpret**

most of people beleive baby was not rude during the flight. only a few people answered very rude.


7. Investigate the relationship between gender and the variables `Do.you.have.any.children.under.18.` and `baby.on.plane`. How is the attitude towards babies on planes shaped by gender and having children under 18? Find a plot that summarises your findings (use `ggplot2`).

```{r}

library(ggmosaic)
ggplot(data = fly)  + 
  geom_mosaic(aes(x = product(Gender), fill=Do.you.have.any.children.under.18., weight=1)) +
  facet_grid(Gender~baby.on.plane.)

```

**Interpretation**

the majority of both females and males that answered baby were rude or somewhat rude on the plane mostly do not have childeren under 18. 

