---
title: 'Stat 480 - Homework #3'
author: "Zahra Khoshmanesh"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Spotify data

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.  

2. Using the command below, read in the spotify data set into your R session. 

```{r}
spotify <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/Stat480-at-ISU.github.io/master/homework/data/spotify.csv")
```

3. Use one of our object inspecting functions and interpret the result in the data that you see.

4. Use the package `ggplot2` to draw a barchart of the genres. In addition, map the genre categories to the fill color of the barchart.

5. Use the package `ggplot2` to draw a histogram of one of the continuous variables in the dataset. Use fill color to show the genre categories and adjust the binwidth if necessary. Use `facet_wrap()` to create a histogram for each of the genre categories.

6. Use the package `ggplot2` to draw a boxplot to compare one of the continuous variables with the genre categories. Again, use fill color to show the genre categories.

7. For each of the three figures above, write a two-three sentence summary, describing the
    a. **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?
    b. **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?
    c. **additional message**: point out anomalies or outliers, if there are any. 


Due date: the homework is due before class on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html file with it. 

## Solution

read data

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.  

2. Using the command below, read in the spotify data set into your R session. 

```{r}
spotify <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/Stat480-at-ISU.github.io/master/homework/data/spotify.csv")
```

3. Use one of our object inspecting functions and interpret the result in the data that you see.

We use summary to know the min,max,median,mean,1st quantile and 3rd quantile of each continous variables and frequencies for categoricals variables in spotify data.From summary we can see that Queen has the highest frequent artist in the data. popularity of each record can be between 0 to 100 with average of 42.78 in the data set.the highest frequent genre is edm following rap. 

```{r}
#use summary to know the min,max,median,mean,1st quantile and 3rd quantile of each variables in spotify data
summary(spotify)
#let us check the structure of the data
str(spotify)


```

```{r}
#print first 6 rows of data
head(spotify)
```


The data is 10000 rows of music track having 13 features such as artist,popularity, subgenre ,...

4. Use the package `ggplot2` to draw a barchart of the genres. In addition, map the genre categories to the fill color of the barchart.

```{r}
library(ggplot2)

ggplot(data=spotify,aes(x=playlist_genre,fill=playlist_genre))+
  geom_bar()

```


**structure of the plot**: this is a barchart. barchart is used when we want to chart a univariate, here **genre**.So, x is **playlist_genre**. we do not have y here. fill is again **playlist_genre** since we want to use different categories in **genre** as map for our chart.

**main message of the plot**: In above barchart we can see the highest frequent of genre and order of different genre. here edm has the highest frequency in the data following rap,pop,r&b and latin. So, edm has the highest frequency. Also all different genre frequency are > 1500 in the data

**additional message**: we can not see any anamoly in barchart. need to use another chart like boxplot to detect outliers.

5. Use the package `ggplot2` to draw a histogram of one of the continuous variables in the dataset. Use fill color to show the genre categories and adjust the binwidth if necessary. Use `facet_wrap()` to create a histogram for each of the genre categories.

**structure of the plot**: this is histogram of popularity for different genre across the data. We selected popularity as x. since histogram is used to sow univariant continious variable, we do not have y. we use genre as filling color and used face_wrap to draw histogram for different categories of genre.

**main message of the plot**: we can see the Symmetry/Skewness in different genre. such as histogram of popularity for edm and rap categoryis symmetric and somehow follows normal distribution or pop and r&b are skewed to left means most of track of these type are less popular. 

**additional message**: we can see huge number of zero popularity in all histograms and maybe indicating some anamolies. 


```{r}

ggplot(data = spotify,aes(x = track_popularity,fill=playlist_genre)) +
  geom_histogram(binwidth = 3) +
  facet_wrap(~playlist_genre) +
  ggtitle("binwidth = 3")


```


6. Use the package `ggplot2` to draw a boxplot to compare one of the continuous variables with the genre categories. Again, use fill color to show the genre categories.

**structure of the plot**: this is boxplot shows  danceability for different genre across the data. We selected playlist_genre as x since want to show boxplot of different genre categories side by side. y is danceability. we use genre as filling color to map the chart .

**main message of the plot**:  boxplot are used for group comparisons and outlier identifications.with boxplot we can compare median, 1quantile,3rd quantile,min and max. rock has the lowest median of dancability among all genre categories and rap has highest median. also we can see that edm,latin,pop and rock are symmetric (mean is close to median) if ignore the outliers.

**additional message**: we can see outliers in all different genre categories.latin has many outliers. rock has only a few outliers.

```{r}

ggplot(data=spotify)+
  geom_boxplot(aes(x =playlist_genre,y=danceability,fill=playlist_genre))
  

```

